### 2013-10-02 chi2014/qualitative.tex
started revision

In addition to our quantitative analysis of the rate of idea generation, we also performed a qualitative analysis.
We selected 30 brainstorming runs (5 from each condition). The primary author applied open coding to identify characteristics of responses and runs that are independent of the uniqueness and originality measures observed in the quantitative section of this paper.
The characteristics primarily took the form of \emph{strategies} employed to different degrees by different participants.
We constructed \emph{personas} around the use of these strategies.

@@

A \emph{strategy} is an identifiable mechanism for idea generation that results in several ideas in a brainstorming run. A key component of a strategy is that it be present in some runs but not in others.

One strategy employed was \emph{feature exploration}. An MP3 player may have many features that distinguish it from any other electronic device of the same physical proportions: audio output (and potentially input), data storage, a screen, general purpose computing abilities (apps), connectivity (WiFi, USB), a camera, a unique appearance, GPS.
Many participants treated the MP3 player as a pocket-sized brick when generating ideas.
In contrast, the feature use strategy led to responses such as:

\begin{itemize}
    \item use them in museums to give information on various installations (audio output)
    \item use in a scavenger hunt, device must be plugged in at another location to see clue
\end{itemize}

Runs utilizing the feature exploration strategy produce responses that seem intuitively better than those that do not. It is satisfying to have a use for a device that "fits" the device as closely as possible, so that none of the functionality is "wasted". Despite this, the proportion of runs utilizing feature exploration seemed roughly constant across all other strategies, and the strategy does not form a basis for discrimination between any personas.

The second strategy identified is \emph{riffing}. While riffing was examined earlier in relation to time spent on idea generation, we found in our qualitative analysis that it was manifested in many ways. \emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:

>>>>

In addition to modelling data and confirming our hypotheses, we performed a qualitative analysis.
We selected 30 brainstorming runs (5 from each condition). The primary author applied open coding to identify characteristics of responses and runs that are independent of the uniqueness and originality measures observed above. In particular, we identified groups of ideas within runs that shared characteristics. These characteristics, in addition to temporal order, form the basis for what we call \emph{strategies}: repeatedly observed classes of ideas that are generated by some participants but not others.
Furthermore, we use the presence or lack of presence of strategies to define \emph{personas}.

@@

A \emph{strategy} is a class of ideas repeatedly observed in groups, though not neccessarily consecutively, within brainstorming runs. These ideas are defined by some common characteristics and may or may not be further defined by temporal order. Not all participants utilized all strategies in their brainstorming runs.

We identified three categories of strategy. The first, \emph{feature exploration}, is specific to the MP3 player problem presented in the paper. This category entails focussing on specific features a device possesses and leveraging them as the basis for generating new ideas. The second category expands on the \emph{riffing} phenomenon introduced earlier, identifying six(?) different types of riffing exployed. Finally, \emph{time-saving} strategies are those that do not neccesarily impact the kinds of ideas generated, but allow the participant to generate at a faster rate.

\subsubsection{Feature exploration}

One category of strategies employed was \emph{feature exploration}. An MP3 player has many features that distinguish it from any other electronic device of the same physical proportions: audio output (and potentially input), data storage, a screen, general purpose computing abilities (apps), connectivity (WiFi, USB), a camera, a unique appearance, GPS.
Each of these features forms the basis for a strategy of idea generation.
Participants would choose a feature and generate ideas that leveraged that feature in various ways. For example, here are several answers from a participant who explored audio input and output:

\begin{itemize}
    \item have them as a resource on public transportation -- people must supply their own headphones
    \item use them in museums to give information on various installations
    \item have cities install them in tourist areas, so people can listen about where they are
\end{itemize}

While it might be expected that all participants would employ this strategy to some degree, many treated the MP3 player as a generic electronic device of equivalent size. Runs were commonly composed of answers that could apply to almost any un-used household item:

\begin{itemize}
    \item doorstop
    \item use in abstract art
    \item recycle
\end{itemize}

Runs that used a feature exploration strategy produce responses that seem intuitively better than those that do not. It is satisfying to have a use for a device that "fits" the device as closely as possible, so that none of the functionality is "wasted".

\subsubsection{Riffing}

The second category of strategies is \emph{riffing}. We found riffing manifested itself many ways, and identified six distinct riffing strategies used by participants: \emph{Specificity riffing} occurs when a participant generates two or more ideas and one is a generalization of the other. For example, two consecutive ideas given by a participant:
